---
title: "winter_datathon_26"
author: "Fiona, Nicolle, Kevin"
date: "2026-02-20"
output: pdf_document
---
#list of cleaned columns
#IMDB rating, Run Time, 

# cleaning dataset

```{r}
library(tidyverse)

x <- c(1)
movies_data <- read.csv("movie_data.csv")
attach(movies_data)
movies_clean <- movies_data
movies_clean %>% map(~sum(is.na(.)))

movies_data <- read.csv("movie_data.csv")
head(movies_data)

movies_data <- drop_na(movies_data)
movies_clean <- movies_data
movies_clean %>% map(~sum(is.na(.)))

movies_clean$Run.Time <- movies_data$Run.Time %>% str_extract_all("\\d+") %>% map(~as.numeric(.)) %>%
  map(~.[1] * 60 + .[2]) %>% unlist()
```

```{r runtime cleanup}
movies_clean$Run.Time <- movies_data$Run.Time %>% str_extract_all("\\d+") %>% map(~as.numeric(.)) %>%
  map(~.[1] * 60 + .[2]) %>% unlist()
```

```{r imdb review cleanup}
km <- function(string){
  if(!is.na(as.numeric(string))){
    num <- as.numeric(string)
  }
  if(str_detect(string, "K")){
    num <- as.numeric(str_sub(string, 1, -2)) * 1000
  }
  if(str_detect(string, "M")){
    num <- as.numeric(str_sub(string, 1, -2)) * 1000000
  }
  num
}

movies_clean$Count.of.IMDb.Ratings <- movies_data$Count.of.IMDb.Ratings %>% 
  map_dbl(~km(.)) 

movies_clean <- drop_na(movies_clean)
```

```{r genre cleanup}
movies_wide_genres <- movies_data %>% 
  dplyr::select(Title, Genre) %>% 
  tidyr::separate_rows(Genre, sep = ",") %>%
  dplyr::mutate(present = TRUE) %>%
  tidyr::pivot_wider(
    names_from = Genre,
    values_from = present,
    values_fill = FALSE,
    values_fn = any
  )

movies_clean <- movies_clean %>% 
  dplyr::inner_join(movies_wide_genres, by = "Title")

```

```{r release date}
library(dplyr)
library(tidyr)

movies_clean <- movies_data |> 
  separate(
    Release.Date,
    into = c("date", "country"),
    sep = "\\(", 
    fill= "right"
    ) |>
        mutate(date = str_trim(date),
                country = str_trim(sub("\\)$", "", country)),
                month = str_extract(date, "^[A-Za-z]+")
              ) |>
                relocate(month, .before = date)
```

```{r budget}
library(tidyverse)
library(readr)
library(stringr)

#assume 3% average annual inflation
base_year <- 2026
annual_inflation <- 0.03 

#identify what currencies are there
movies_data %>%
  mutate(budget_currency = str_trim(str_extract(Budget, "^[^0-9]+"))) %>%
  count(budget_currency, sort = TRUE)

#exchange rates as of 20 Feb 2026
exchange_rates <- tibble(
  budget_currency = c("$", "£", "EUR", "CA$", "A$", "¥", "RUR", "₹", "CN¥", "R$", "THB"),
  rate_to_usd = c(
    1.00,   # USD
    1.35,   # GBP
    1.18,   # EUR
    0.73,   # CAD
    0.71,   # AUD
    0.0065, # JPY
    0.013,  # RUR (old russian ruble)
    0.011,  # INR
    0.14,   # CNY
    0.19,   # BRL
    0.032   # THB
  )
)

#convert
movies_clean <- movies_data %>%
  mutate(
    budget_usd_2026 = {
      currency <- str_trim(str_extract(Budget, "^[^0-9]+"))
      currency[is.na(currency)] <- "$"

      # standardize: old currencies to euro
      currency[currency == "€"] <- "EUR"
      currency[currency %in% c("DEM", "FRF")] <- "EUR"

      value <- parse_number(Budget) 

      rate <- exchange_rates$rate_to_usd[match(currency, exchange_rates$budget_currency)]
      rate[is.na(rate)] <- 1  # if no currency, treat as USD

      years <- base_year - as.integer(Year) 

      (value * rate) * (1 + annual_inflation)^years #inflate to 2026 USD (future value formula)
    }
  )
```

```{r domestic gross}
class(movies_clean[["Domestic.Gross"]][1]) #so we know its a character

movies_clean[["Domestic.Gross"]] <- as.numeric(gsub("[\\$,]", "", movies_clean[["Domestic.Gross"]]))

class(movies_clean[["Domestic.Gross"]][1])

#apply for other 2 gross figs
```

```{r domestic weekend gross}
movies_clean[["Domestic.Weekend.Gross"]] <- as.numeric(gsub("[\\$,]", "", movies_clean[["Domestic.Weekend.Gross"]]))
```

```{r worldwide gross}
movies_clean[["Worldwide.Gross"]] <- as.numeric(gsub("[\\$,]", "", movies_clean[["Worldwide.Gross"]]))
```

```{r ratings cleanup}
movies_clean[["Rotten.Tomatoes.Rating"]] <- as.numeric(gsub("%", "", movies_clean[["Rotten.Tomatoes.Rating"]]))
```


```{r gross account for inflation}
library(dplyr)
library(readr)

movies_clean <- movies_clean %>%
  mutate(
    domestic_gross_2026 =
      Domestic.Gross *
      (1 + annual_inflation)^(base_year - as.integer(Year)),

    domestic_weekend_gross_2026 =
      Domestic.Weekend.Gross *
      (1 + annual_inflation)^(base_year - as.integer(Year)),

    worldwide_gross_2026 =
      Worldwide.Gross *
      (1 + annual_inflation)^(base_year - as.integer(Year))
  )
```

```{r}
#clean plot
movies_clean <- movies_clean %>%
  mutate(
    Plot = str_replace_all(Plot, "&apos;", "'"),
    Plot = str_replace_all(Plot, "&quot;", "\""),
    Plot = str_replace_all(Plot, "&amp;", "&")
  )
```

```{r}
#install.packages("tidytext")
#install.packages("textdata")
library(tidytext)
library(textdata)
library(dplyr)
library(stringr)

plot_words <- movies_clean %>%
  select(Title, Plot) %>%
  mutate(Plot = tolower(Plot)) %>%
  unnest_tokens(word, Plot)

afinn <- get_sentiments("afinn")

plot_sentiment <- plot_words %>%
  inner_join(afinn, by = "word") %>%
  group_by(Title) %>%
  summarise(
    plot_sentiment = mean(value),
    sentiment_sum = sum(value),
    word_count = n()
  )

movies_clean <- movies_clean %>%
  left_join(plot_sentiment, by = "Title")


```

```{r}
library(tidytext)
library(dplyr)

plot_words <- movies_clean %>%
  unnest_tokens(word, Plot)

plot_words %>%
  count(word, sort = TRUE) %>%
  filter(!word %in% stop_words$word) %>%
  slice_max(n, n = 100)

```
```{r}
library(ggplot2)
library(tidytext)
library(dplyr)

plot_words <- movies_clean %>%
  unnest_tokens(word, Plot)

top_words <- plot_words %>%
  count(word, sort = TRUE) %>%
  filter(!word %in% stop_words$word) %>%
  slice_max(n, n = 100)

ggplot(top_words,
       aes(x = reorder(word, n),
           y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    x = "Word",
    y = "Frequency",
    title = "Top 100 Most Frequent Words in Movie Plots"
  ) +
  theme_minimal()
```
```{r}
library(tidytext)
library(dplyr)
library(ggplot2)

plot_words <- movies_clean %>%
  unnest_tokens(word, Plot)

word_ratings <- plot_words %>%
  filter(!word %in% stop_words$word) %>%
  group_by(word) %>%
  summarize(
    avg_rating = mean(IMDb.Rating, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n >= 20) %>%  
  slice_max(n, n = 30)   

ggplot(word_ratings,
       aes(x = reorder(word, avg_rating),
           y = avg_rating)) +
  geom_col() +
  coord_flip() +
  labs(
    x = "Word",
    y = "Average IMDb Rating",
    title = "Average Rating of Movies Containing Each Word"
  ) +
  theme_minimal()
```
```{r}

```


```{r}
model <- lm(world_gross_2026 ~ 
     plot_sentiment +
     budget_usd_2026 +
     IMDb.Rating,
   data = movies_clean)

summary(model)
plot(model)
```

```{r}
#runtime effect on ratings 

#rotten tomato ratings
attach(movies_clean)
runtime_vs_rating_rt <- lm(Rotten.Tomatoes.Rating ~ Run.Time)
summary(runtime_vs_rating_rt)
plot(runtime_vs_rating_rt)

```

```{r}
#IMdb ratings
attach(movies_clean)
imdb_ratings_proportion <- IMDb.Rating * Count.of.IMDb.Ratings
runtime_vs_rating_imdb <- lm(imdb_ratings_proportion ~ Run.Time)
summary(runtime_vs_rating_imdb)
plot(runtime_vs_rating_imdb)
```


```{r}
#metascore ratings
runtime_vs_rating_meta <- lm(Metascore.Rating ~ Run.Time)
summary(runtime_vs_rating_meta)
plot(runtime_vs_rating_meta)
```

```{r}
library(ggplot2)
library(dplyr)

ratings_long <- movies_clean %>%
  dplyr::select(
    Run.Time,
    Rotten.Tomatoes.Rating,
    Metascore.Rating
  ) %>%
  pivot_longer(
    -Run.Time,
    names_to = "Rating_Type",
    values_to = "Rating"
  )

ggplot(ratings_long,
       aes(x = Run.Time,
           y = Rating,
           color = Rating_Type)) +
  geom_point(alpha = 0.6) +
  ylim(0, 100)
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Runtime (minutes)",
    y = "Rating",
    title = "Relationship Between Runtime and Movie Ratings"
  ) +
  theme_minimal()
```
```{r}

ggplot(movies_clean, aes(Run.Time, imdb_ratings_proportion)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Relationship between Runtime and IMdb Ratings")
```
```{r}
q25 <- quantile(movies_clean$IMDb.Rating, 0.25, na.rm = TRUE)

low_movies <- movies_clean %>%
  filter(IMDb.Rating <= q25)

summary(low_movies$Run.Time)

q25 <- quantile(movies_clean$Rotten.Tomatoes.Rating, 0.25, na.rm = TRUE)

low_movies <- movies_clean %>%
  filter(Rotten.Tomatoes.Rating <= q25)

summary(low_movies$Run.Time)

q25 <- quantile(movies_clean$Metascore.Rating, 0.25, na.rm = TRUE)

low_movies <- movies_clean %>%
  filter(Metascore.Rating <= q25)

summary(low_movies$Run.Time)
```
```{r}
library(ggplot2)
library(dplyr)

ratings_long <- movies_clean %>%
  dplyr::select(
    Release.Month,
    Rotten.Tomatoes.Rating,
    Metascore.Rating
  ) %>%
  pivot_longer(
    -Run.Time,
    names_to = "Rating_Type",
    values_to = "Rating"
  )

ggplot(ratings_long,
       aes(x = Run.Time,
           y = Rating,
           color = Rating_Type)) +
  geom_point(alpha = 0.6) +
  ylim(0, 100)
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Runtime (minutes)",
    y = "Rating",
    title = "Relationship Between Runtime and Movie Ratings"
  ) +
  theme_minimal()
```


```{r}
#clean tagline
library(dplyr)
library(stringr)
library(tidytext)
library(tidyr)

na_vals <- c("NA", "N/A", "NULL", "na", "n/a", "null")

movies_clean <- movies_clean %>%
  mutate(
    Tagline = str_squish(Tagline),
    Tagline = str_replace_all(Tagline, "&apos;", "'"),
    Tagline = ifelse(Tagline %in% na_vals, NA_character_, Tagline)
  )

#for text mining: so all lower case to make it easy 
tag_base <- movies_clean %>%
  select(Title, Tagline) %>%
  filter(!is.na(Tagline), Tagline != "")
```

```{r}
#check for common tagline words 
tag_words <- tag_base %>%
  mutate(Tagline_lower = str_to_lower(Tagline)) %>%
  unnest_tokens(word, Tagline_lower) %>%
  anti_join(stop_words, by = "word")

tag_words %>% # one word
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20)

tag_bigrams <- tag_base %>% #two words with word fillers "the, of" gone
  mutate(Tagline_lower = str_to_lower(Tagline)) %>%
  unnest_tokens(bigram, Tagline_lower, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word
  ) %>%
  unite(bigram, word1, word2, sep = " ")

tag_bigrams %>%
  count(bigram, sort = TRUE) %>%
  slice_max(n, n = 20)

tag_trigrams <- tag_base %>% #three words with word fillers gone
  mutate(Tagline_lower = str_to_lower(Tagline)) %>%
  unnest_tokens(trigram, Tagline_lower, token = "ngrams", n = 3) %>%
  separate(trigram, into = c("word1", "word2", "word3"), sep = " ") %>%
  filter(
    !word1 %in% stop_words$word,
    !word2 %in% stop_words$word,
    !word3 %in% stop_words$word
  ) %>%
  unite(trigram, word1, word2, word3, sep = " ")

tag_trigrams %>%
  count(trigram, sort = TRUE) %>%
  slice_max(n, n = 20)

```

```{r}
#check tagline styling features:

#check word count/length
tagline_word_count <- str_count(movies_clean$Tagline, "\\S+")
tagline_char_length <- nchar(movies_clean$Tagline)

exclaim_count <- str_count(movies_clean$Tagline, "!")
question_count <- str_count(movies_clean$Tagline, "\\?")
comma_count <- str_count(movies_clean$Tagline, ",")
dot_count <- str_count(movies_clean$Tagline, "\\.")

has_exclaim <- as.numeric(exclaim_count > 0)
has_question <- as.numeric(question_count > 0)

# check if capitalization matter
upper_chars <- str_count(movies_clean$Tagline, "[A-Z]")
letters_only <- str_count(movies_clean$Tagline, "[A-Za-z]")
upper_ratio <- ifelse(letters_only == 0, 0, upper_chars / letters_only)
all_caps <- as.numeric(letters_only > 0 & upper_ratio > 0.9)

# 2) check word repetition matter or no
tag_rep <- movies_clean %>%
  select(Title, Tagline) %>%
  filter(!is.na(Tagline), Tagline != "") %>%
  mutate(Tagline_lower = str_to_lower(Tagline)) %>%
  unnest_tokens(word, Tagline_lower) %>%
  group_by(Title) %>%
  summarise(repetition_ratio = 1 - (n_distinct(word) / n()), .groups = "drop")

repetition_ratio <- tag_rep$repetition_ratio[match(movies_clean$Title, tag_rep$Title)]
repetition_ratio[is.na(repetition_ratio)] <- 0

# 3) model: do tagline style choices predict revenue?
m_tagline <- lm(
  worldwide_gross_2026 ~
    tagline_word_count +
    has_exclaim +
    has_question +
    upper_ratio +
    all_caps +
    repetition_ratio +
    budget_usd_2026 +
    IMDb.Rating,
  data = movies_clean
)

summary(m_tagline)
```
Each additional word in the tagline is associated with ~ $3.2 million more revenue, holding other variables constant. Statistically significant (p < 0.05). Longer taglines slightly outperform shorter ones. The rest don't matter.

```{r}
log_revenue <- log(movies_clean$worldwide_gross_2026 + 1) #revenue skewed so to make it fair 

m_tagline_log <- lm( 
  log_revenue ~
    tagline_word_count +
    has_exclaim +
    has_question +
    upper_ratio +
    all_caps +
    repetition_ratio +
    budget_usd_2026 +
    IMDb.Rating,
  data = movies_clean
)

summary(m_tagline_log)
```
Earlier "longer taglines increase revenue" effect dont matter after log-transform. Means the previous significance was likely driven by extreme blockbuster scale effects. Movies with exclamation marks in their tagline earn roughly 25–30% less revenue, controlling for budget and rating. Not strong enough for p < 0.05, but kinda. Budget Matters. Each 1-point increase in IMDb rating leads to approx. 32% increase in revenue.


