---
title: "winter_datathon_26"
author: "Fiona, Nicolle, Kevin"
date: "2026-02-20"
output: pdf_document
---
#list of cleaned columns
#IMDB rating, Run Time, 

# cleaning dataset

```{r}
library(tidyverse)

x <- c(1)
movies <- read.csv("movie_data.csv")
attach(movies)
movies_clean <- movies
movies %>% map(~sum(is.na(.)))

movies_data <- read.csv("movie_data.csv")
head(movies_data)

movies_data <- drop_na(movies_data)
movies_clean <- movies_data
movies_clean %>% map(~sum(is.na(.)))

movies_clean$Run.Time <- movies$Run.Time %>% str_extract_all("\\d+") %>% map(~as.numeric(.)) %>%
  map(~.[1] * 60 + .[2]) %>% unlist()
```

```{r runtime cleanup}
movies_clean$Run.Time <- movies_data$Run.Time %>% str_extract_all("\\d+") %>% map(~as.numeric(.)) %>%
  map(~.[1] * 60 + .[2]) %>% unlist()
```

```{r imdb review cleanup}
km <- function(string){
  if(!is.na(as.numeric(string))){
    num <- as.numeric(string)
  }
  if(str_detect(string, "K")){
    num <- as.numeric(str_sub(string, 1, -2)) * 1000
  }
  if(str_detect(string, "M")){
    num <- as.numeric(str_sub(string, 1, -2)) * 1000000
  }
  num
}

movies_clean$Count.of.IMDb.Ratings <- movies_data$Count.of.IMDb.Ratings %>% 
  map_dbl(~km(.)) 
```

```{r genre cleanup}
movies_wide <- movies_data %>% select(Title, Genre) %>% 
  separate_rows(Genre, sep = ",") %>%
  mutate(present = TRUE) %>%
  pivot_wider(
    names_from = Genre,
    values_from = present,
    values_fill = FALSE,
    values_fn = any
  )

movies_clean <- movies_clean %>% inner_join(movies_wide, by = "Title")
```

<<<<<<< HEAD
```{r release date}
=======
```{r relsease date}
library(dplyr)
library(tidyr)
>>>>>>> 49202af4bdc29413dc58dba63dea833cfe13e7eb

movies_data |> 
  separate(
    Release.Date,
    into = c("date", "country"),
    sep = "\\(", 
    fill= "right"
    ) |>
        mutate(date = str_trim(date),
                country = str_trim(sub("\\)$", "", country)),
                month = str_extract(date, "^[A-Za-z]+")
              ) |>
                relocate(month, .before = date)
```

```{r budget}

```

<<<<<<< HEAD
```{r budget}
library(tidyverse)
library(readr)
library(stringr)

#assume 3% average annual inflation
base_year <- 2026
annual_inflation <- 0.03 

#identify what currencies are there
movies_data %>%
  mutate(budget_currency = str_trim(str_extract(Budget, "^[^0-9]+"))) %>%
  count(budget_currency, sort = TRUE)

#exchange rates as of 20 Feb 2026
exchange_rates <- tibble(
  budget_currency = c("$", "£", "EUR", "CA$", "A$", "¥", "RUR", "₹", "CN¥", "R$", "THB"),
  rate_to_usd = c(
    1.00,   # USD
    1.35,   # GBP
    1.18,   # EUR
    0.73,   # CAD
    0.71,   # AUD
    0.0065, # JPY
    0.013,  # RUR (old russian ruble)
    0.011,  # INR
    0.14,   # CNY
    0.19,   # BRL
    0.032   # THB
  )
)

#convert
movies_clean <- movies_data %>%
  mutate(
    budget_usd_2026 = {
      currency <- str_trim(str_extract(Budget, "^[^0-9]+"))
      currency[is.na(currency)] <- "$"

      # standardize: old currencies to euro
      currency[currency == "€"] <- "EUR"
      currency[currency %in% c("DEM", "FRF")] <- "EUR"

      value <- parse_number(Budget) 

      rate <- exchange_rates$rate_to_usd[match(currency, exchange_rates$budget_currency)]
      rate[is.na(rate)] <- 1  # if no currency, treat as USD

      years <- base_year - as.integer(Year) 

      (value * rate) * (1 + annual_inflation)^years #inflate to 2026 USD (future value formula)
    }
  )
```

=======
```{r domestic gross}
class(movies[["Domestic.Gross"]][1]) #so we know its a character

movies["Domestic.Gross"] <- as.numeric(movies[["Domesic.Gross"]])

class(movies[["Domestic.Gross"]][1])

```
```{r domestic weekend gross}
movies[["Domestic.Weekend.Gross"]] <- as.numeric(str_sub("[\\$,]", "", movies[["Domestic.Weekend.Gross"]]))

class(movies[["Domestic.Weekend.Gross"]][1])
```

```{r}

```


>>>>>>> 49202af4bdc29413dc58dba63dea833cfe13e7eb
